\begin{tabularx}{\textwidth}{lX}
%\begin{tabular}{lp{8cm}}
  \toprule
  Source & Die Armee teilte zugleich mit, dass in den vergangenen 24
           Stunden sieben Soldaten getötet worden seien.  \\
  \addlinespace
  Reference & The army announced at the same time that over the
              past 24 hours,
              seven soldiers have been killed. \\
  \addlinespace
  \acs{ar} & The army also announced that seven soldiers have been killed over
             the past 24 hours.\\
  \addlinespace
  \acs{nar} & The army also announced that seven soldiers \redund{had}
              \redund{killed} killed in the last \redund{24} 24 hours. \\
         & \quad \it Missing verb, Repeated tokens \\
  \addlinespace
  \acs{nar} + \acs{lm}
         & The army also announced that \redund{the} seven soldiers
           \greenund{had been killed} in the last \greenund{24} hours.  \\
         & \quad \it Fixed issues but inserted a definite article \\

  \midrule

  Source & Wir waren früher ein freundliches Land, wie kann es sein, dass wir
           nun kämpfen? \\
  \addlinespace

  Reference & We used to be one friendly country in the past, how did it happen
              that we are fighting now? \\
  \addlinespace

  \acs{ar} & We used to be a friendly country, how can we fight now? \\
         & \quad \it Translation too short \\
  \addlinespace

  \acs{nar}
         & We \redund{were} a friendly country, \redund{how can it we}
              are fighting now? \\
         & \quad \it Missing words \\
  \addlinespace

  \acs{nar} + \acs{lm}
         & We \redund{were} a friendly country, \greenund{how can it be}
           that we are fighting now? \\
         & \quad \it Partially correct \\

  % \midrule

  % Source & Ein 78-jähriger Mann steht wegen des Überfahrens eines
  %          dreijährigen Mädchens auf einem Fußgängerüberweg in Fife
  %          vor Gericht. \\

  % \addlinespace
  % Reference  & A 78-year-old man is to stand trial accused of running
  %              over a three-year-old girl on a pelican crossing in Fife. \\

  % \addlinespace
  % \acs{ar} &  A 78-year-old man is in court for crossing a three-year-old girl on a pedestrian crossing in Fife. \\

  % \addlinespace
  % \acs{nar} &  A 78-year-old man isold girl on a pedestrian in Fifeyear girl. \\

  % \addlinespace
  % \acs{nar} + \acs{lm} \\
  % A 78-year-old man is on a pedestrian crossing in Fife.


  \midrule

  Source & Seiner Ansicht nach könnten alle Mitglieder beider Vereine künftig wieder an einem Strang ziehen. \\
  \addlinespace

  Reference & In his opinion, all members of each club could come together again in the future. \\
  \addlinespace

  \acs{ar} & In his view, all members of both clubs could pull together again in the future. \\
  \addlinespace

  \acs{nar} & \redund{In} believes that all members of both clubs could pull \redund{together future}. \\
         & \quad \it Ungrammatical beginning and end \\
  \addlinespace

  \acs{nar} + \acs{lm}
         & \greenund{He} believes that all \blueund{the} members of both
           clubs could pull together. \\
         & \quad \it Fluency corrections; inserted ``the'' again, this time
           not a mistake \\


 % \midrule
  % Source & \\
  % Reference & \\
  % \acs{ar} & \\
  % \acs{nar} & \\
  % \acs{nar} + \acs{lm} & \\

  \bottomrule
%\end{tabular}
\end{tabularx}

%    nAR &  \blueund{Aggressivitätivität}  \\
%    & \quad\rotatebox[origin=c]{180}{$\Lsh$} \it  with repeated subwords (blue). \\

%    nAR + LM &  \greenund{angeborenen}  \\
%    & \quad\rotatebox[origin=c]{180}{$\Lsh$} \it  chosen. \\

%    AR &  \greenund{angeborenen} . \\

%    Reference & Aufgrund ihrer ureigenen Aggressivität wurden Songs dieser Art nicht
% mehr auf der Konsole gespielt. \\
%   \midrule


%   \bottomrule
% \end{tabular}